---
title: "Week 2 live lecture notes"
author: "Garth Tarr"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sway quiz

https://edstem.org/au/courses/6600/sway/

# Hypothesis testing review

[Handwritten notes](hyp_test.pdf)

# Odds ratios in the wild

- Simoes et al. (2021). Effects of pre-operative isolation on postoperative pulmonary complications after elective surgery: an international prospective cohort study. _Anaesthesia._ DOI: [10.1111/anae.15560](https://associationofanaesthetists-publications.onlinelibrary.wiley.com/doi/full/10.1111/anae.15560)

```{r}
table2 = matrix(c(67993, 1386,
           6877, 91,
           10483, 206,
           8910, 252), 
           ncol = 2, byrow = TRUE)
colnames(table2) = c("No complications", "Postop complication(s)")
rownames(table2) = c("No isolation","<= 3 days", "4-7 days",">= 8 days")
table2
```

## <= 3 days isolation

Relative to baseline

```{r}
e1 = table2[1:2,2:1]
e1
# install.packages("mosaic")
mosaic::oddsRatio(e1, verbose = TRUE)
```

Understanding this output:


```{r}
e1
rowsums = rowSums(e1)
# looking within row 1 (no isolation)
p1 = e1[1, 1]/rowsums[1]
# looking within row 1 (<= 3 days isolation)
p2 = e1[2, 1]/rowsums[2]
o1 = p1/(1 - p1)
o2 = p2/(1 - p2)
RR = p2/p1
OR = o2/o1
RR
OR
```

So if we're interpreting this, 

```{r}
# OR confidence interval
se = sqrt(sum(1/e1))
log_or_ci = log(OR) + c(-1,1)*qnorm(0.95)*se
or_ci = exp(log_or_ci)
or_ci
```

Is this a statistically significant result? At what level of significance?




## 4-7 days isolation

Relative to baseline

```{r}
e2 = table2[c(1,3),2:1]
e2
# install.packages("mosaic")
mosaic::oddsRatio(e2)
mosaic::relrisk(e2)
```


## >= 8 days isolation

Relative to baseline

```{r}
e3 = table2[c(1,4),2:1]
e3
mosaic::oddsRatio(e3)
mosaic::relrisk(e3)
```

