```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Smoking data

Here's the smoking data (before it is broken down by age group).

```{r, message = FALSE}
x = read_csv("https://git.io/fNGXk")
# in case of catastrophic failure (internet not working):
# x = tibble::tribble(
#        ~status,  ~survival, ~count,
#       "Smoker",     "Died",    139,
#       "Smoker", "Survived",    443,
#   "Non-smoker",     "Died",    230,
#   "Non-smoker", "Survived",    502
#   )
```

Convert into long form.

```{r}
x_long = tidyr::uncount(x, weights = count)
```

### Calculate summary statistics

Calculate group level mortality rates.  Include a sample size column.  Think about an alternative to `sum(...)/n()`.

```{r}
x_long %>%
  dplyr::group_by(status) %>% 
  dplyr::summarise(
    rate = sum(survival == "Died")/n()
  )
```

Calculate overall mortality rate.

```{r}
x_long %>%
  dplyr::summarise(
    rate = sum(survival == "Died")/n()
  )
```

### Plots

Stacked bar chart showing mortality rates for smokers and non-smokers.

```{r}
ggplot(x, aes(x = status, y = count, fill = survival)) + 
  geom_bar(stat = "identity")

# compare with:
ggplot(x_long, aes(x = status, fill = survival)) + 
  geom_bar()
```


## Smoking data by age group

```{r, message = FALSE}
y = readr::read_csv("https://git.io/fNGSi")
# in case of catastrophic failure:
# y = tibble::tribble(
#        ~status,  ~survival, ~age_group, ~count,
#       "Smoker",     "Died",    "18-24",      2,
#       "Smoker",     "Died",    "25-34",      3,
#       "Smoker",     "Died",    "35-44",     14,
#       "Smoker",     "Died",    "45-54",     27,
#       "Smoker",     "Died",    "55-64",     51,
#       "Smoker",     "Died",    "65-74",     29,
#       "Smoker",     "Died",      "75+",     13,
#       "Smoker", "Survived",    "18-24",     53,
#       "Smoker", "Survived",    "25-34",    121,
#       "Smoker", "Survived",    "35-44",     95,
#       "Smoker", "Survived",    "45-54",    103,
#       "Smoker", "Survived",    "55-64",     64,
#       "Smoker", "Survived",    "65-74",      7,
#       "Smoker", "Survived",      "75+",      0,
#   "Non-smoker",     "Died",    "18-24",      1,
#   "Non-smoker",     "Died",    "25-34",      5,
#   "Non-smoker",     "Died",    "35-44",      7,
#   "Non-smoker",     "Died",    "45-54",     12,
#   "Non-smoker",     "Died",    "55-64",     40,
#   "Non-smoker",     "Died",    "65-74",    101,
#   "Non-smoker",     "Died",      "75+",     64,
#   "Non-smoker", "Survived",    "18-24",     61,
#   "Non-smoker", "Survived",    "25-34",    152,
#   "Non-smoker", "Survived",    "35-44",    114,
#   "Non-smoker", "Survived",    "45-54",     66,
#   "Non-smoker", "Survived",    "55-64",     81,
#   "Non-smoker", "Survived",    "65-74",     28,
#   "Non-smoker", "Survived",      "75+",      0
#   )
dplyr::glimpse(y)
```

### Cross tabulation

```{r}
# might need dev version of tidyr for pivot_wider
# devtools::install_github("tidyverse/tidyr") 
ytab = y %>%
  tidyr::pivot_wider(id_cols = age_group, 
                     names_from = c(status, survival), 
                     values_from = count,
                     names_sep = ": ")
ytab %>% rename(`Age group` = age_group) %>% knitr::kable()
```

### Calculate mortality rates

```{r}
mortality = y %>% 
  uncount(weights = count) %>% 
  group_by(status, age_group) %>% 
  summarise(rate = mean(survival=="Died"))
```

### Visualise mortality rates

```{r}
mortality %>% 
  ggplot() + 
  aes(x = age_group, y = rate, fill = status) + 
  geom_bar(stat = "identity")
```

### Visualise sample sizes

```{r}
# convert into "long" form
y_long = y %>%  uncount(weights = count)

y_long %>% 
  ggplot() + 
  aes(x = age_group, fill = status) + 
  geom_bar()

# compare with:
y %>% 
  ggplot() + 
  aes(x = age_group, y = count, fill = status) + 
  geom_bar(stat = "identity")
```

## Iris

```{r}
clean_iris = iris %>% janitor::clean_names()
glimpse(clean_iris)
cor(clean_iris$sepal_length, clean_iris$sepal_width)
# the dplyr method
# clean_iris %>% 
#   summarise(
#     Correlation = cor(sepal_length, sepal_width)
#   )
```

BUT when we break it down by species:

```{r}
clean_iris %>% 
  group_by(species) %>% 
  summarise(
    Correlation = cor(sepal_length, sepal_width)
  )
```

Visually what's going on here?

```{r}
clean_iris %>% 
  ggplot() + 
  aes(x = sepal_length, y = sepal_width) 
```

