---
title: "lab 3"
author: "Mason Wong"
date: "16th March 2022"
output: pdf_document
---
```{r setup, warning=F, message=F, include = FALSE}
library(tidyverse)
library(dplyr)
# library(broom)
# library(patchwork)
knitr::opts_chunk$set(echo = TRUE, tidy = FALSE, fig.align = "center", fig.pos = "H")
```

# Question 1

(a)

```{r}
olympic = read.table("~/Desktop/R-programming/3002-labs/lab-2/olympic.txt", sep = "\t", header = TRUE)
head(olympic, 6)
summary(olympic)
tail(olympic, 6)

convert_inch_to_m = function(x) {
  return(x/39.3701)
}
olympicMetric = olympic %>% 
  mutate(across(c(HighJump, DiscusThrow, LongJump), convert_inch_to_m)) %>% 
  mutate(Year = Year + 1900)
```

(b)
```{r}
olympicLm = lm(HighJump ~ LongJump, data = olympicMetric)
SSE_H1 = sum((olympicLm$residuals)^2)
df_H1 = olympicLm$df.residual
summary(olympicLm)
```
(c)

- Under $H_1$ we don't have constraint on $\beta_0$ so we have the full model of $y_i = \beta_0 + \beta_1 x_i + \varepsilon_i$ for $i = 1, 2, 3..., n$ where $y_i$ is the `HighJump` variable and $x_i$ is the `LongJump` variable. 

- Under $H_0$ we do have the constrain that $\beta = 0$ so we have the restricted model is $y_i = \beta_1x_i + \varepsilon_i$ for $i = 1, 2, 3..., n$

(d) 
```{r}
olympic_H0 = lm(HighJump ~ -1 + LongJump, data = olympicMetric)
SSE_H0 = sum((olympic_H0$residuals)^2)
df_H0 = olympic_H0$df.residual
summary(olympic_H0)
```

(e)

```{r, include = FALSE}
f0 = ((SSE_H0 - SSE_H1)/(df_H0 - df_H1))/(SSE_H1/df_H1)
p_val = pf(q = f0, df1 = df_H0 - df_H1, df2 = df_H1, lower.tail = FALSE)
p_val
names(anova(olympic_H0, olympicLm))
anova(olympic_H0, olympicLm)$'Pr(>F)'[2]
```
The observed f statistic is `r f0`. The corresponding p value is `r p_val` which is the same as if we did the command `anova(olympic_H0, olympicLm)$'Pr(>F)'[2]`

(f)
We have that: 

- we are testing the intercept here: 

$$H_0: \beta_0 = 0 \quad\quad vs\quad\quad H_1: \beta_0 \neq 0$$

- The manually computed f statistic for this test was: 

$$`r f0`$$

- The t-test test statistic, testing the same test for the intercept $\beta_0$ was: 

$$0.311$$

- We note that $(0.311)^2 = `r f0`$
```{r}

```


